---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# RcppRustCompare

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
![R-CMD-check](https://github.com/tjtnew/RcppRustCompare/workflows/R-CMD-check/badge.svg?branch=master)
<!-- badges: end -->

Quick comparison of Rust/Rcpp integration in R vis-a-vis the generation of 
Erdos-Renyi random graphs (adjacency lists).

## Rust installation
Visit https://www.rust-lang.org/tools/install for information on installing 
Rust.  Note that `cargo` is only needed at build-time.

## Package installation
`devtools::install_github("tjtnew/RcppRustCompare")`

## Example
```{r, eval=TRUE}
library(RcppRustCompare)
library(ggraph)
library(tidygraph)

# small graphs
n <- 50 # size of graph
p <- 0.2 # probability of edge between two nodes

ggraph(as_tbl_graph(erdos_rust(n, p))) +
  geom_edge_link() +
  geom_node_point()

ggraph(as_tbl_graph(erdos_rcpp(n, p))) +
  geom_edge_link() +
  geom_node_point()


## larger graphs - we check average degree is roughly p * (n - 1)
n <- 10000
avk <- 5
p <- avk / (n - 1)
graph_rust <- erdos_rust(n, p)
graph_rcpp <- erdos_rcpp(n, p)

mean(sapply(graph_rust, length))
mean(sapply(graph_rcpp, length))
```

## benchmarks - dense graphs
Comparable performance between the Rcpp implementation and Rust.
```{r, eval=TRUE}
library(microbenchmark)

microbenchmark(
  erdos_rust(10000, 0.1),
  erdos_rcpp(10000, 0.1),
  erdos_rust(10000, 0.2),
  erdos_rcpp(10000, 0.2),
  erdos_rust(10000, 0.4),
  erdos_rcpp(10000, 0.4),
  erdos_rust(10000, 0.8),
  erdos_rcpp(10000, 0.8),
  times = 10
)
```

## benchmarks - sparse graph
For sparse graphs we can use the algorithm from Batagelj and Brandes, 
[Efficient Generation of Large Random Networks](https://journals.aps.org/pre/abstract/10.1103/PhysRevE.71.036113).

```{r}
n <- 1000000
avk <- 5
p <- avk / (n - 1)
microbenchmark(fast_erdos_rust(n, p), times = 10)
```

This even compares well with the implementation in igraph (although igraph
has a far more complete graph representation than this toy code):

```{r}
library(igraph)
microbenchmark(sample_gnp(n, p), times = 10)
```

